---
title: "Reproducible Research: Peer Assessment 1"
author: "gspinnada uhu"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Unzipping & Loading data
```{r}
unzip("./activity.zip")
activity<-read.csv("./activity.csv")
```
Reclassing date as Date and steps as numeric
```{r}
activity$date<-as.Date(activity$date,"%Y-%m-%d")
activity$steps<-as.numeric(activity$steps)
```


## What is mean total number of steps taken per day?
1.  Calculating the total number of steps taken per day
    ```{r}
    totalsteps_day<-with(activity,tapply(steps,date,sum,na.rm=TRUE))
    ```
2.  Making a histogram of the total number of steps taken each day
    ```{r fig.align='center'}
    hist(totalsteps_day,
      main="Total Number of Steps taken per Day",
      xlab="Steps")
    ```
3.  Calculating the mean and median of the total number of steps taken per day
  a) Mean
    ```{r}
    mean(totalsteps_day)
    ```
  b) Median
    ```{r}
    median(totalsteps_day)
    ```


## What is the average daily activity pattern?
1.  Time series plot of the 5-minute intervals in a day (x) vs the average number of steps taken per interval for all days (y)
  a) Calculating average number of steps per interval + creating dataframe with averages and interval numbers
    ```{r}
    mean_steps_int<-with(activity,tapply(steps,interval,mean,na.rm=TRUE))
    unique_int<-unique(activity$interval)
    unique_df<-data.frame(unique_int,mean_steps_int)
    names(unique_df)<-c("unique_int","mean_steps")
    ```
  b) Creating plot, using unique values in interval as x-axis and mean_steps_int as y
    ```{r}
    library(ggplot2)
    ggplot(data=unique_df,aes(x=unique_int,y=mean_steps))+
            geom_point()+
            labs(x="Interval Number",y="Average Steps",title="Average Steps per Interval")
    ```

2.  Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
    ```{r}
    library(dplyr)
    int_max_step_avg<-filter(unique_df,mean_steps==max(mean_steps))
    ```
    Interval number `r int_max_step_avg[1,1]` has the highest average number of steps - `r int_max_step_avg[1,2]`.

## Imputing missing values
1.  Counting the total number of rows containing any NA values



## Are there differences in activity patterns between weekdays and weekends?